---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
if(!file.exists('activity.csv')){
  unzip("activity.zip")
}
activity <- read.csv('activity.csv', header=TRUE)
```

From a previous assignment I have learned it's best to turn the date column into a POSIXlt datetime column for ease of use in graphing.

```{r}
dates <- as.character(activity$date)
times <- activity$interval
times <- sprintf("%02d", times)
times <- paste("00", times, sep=":")

activity$datetime <- as.POSIXlt(paste(dates, times), format="%Y-%m-%d %H:%M", tz="GMT")
rm(dates, times)

```



To get the axis to format nicely I added the following to my ggplot2 commands:

+ scale_x_chron(format="%H:%M") 


## What is mean total number of steps taken per day?
First we calculate the total steps per day using the summarize command of the dplyr package.  (We must load plyr first though.)

```{r}
require(plyr)
require(dplyr)
steps_per_day <- summarise(group_by(activity, date), sum(steps))
```

Next we create the histogram:

```{r}

```
## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
